---
title: "Relational databases"
author: "Youri Lam"
date: "`r Sys.Date()`"
output: html_document
---
# Relational data and databases

## Introduction {-}

This assignment was focused on learning the basics of the SQL language and how to work this around relational data and databases.

----

## DBeaver connection {-}

I started by creating three data frames out of files and making them tidy.

```{r import and tidy}
## Import, first 11 lines are metadata
flu_data <- read_csv("raw_data/flu_data.csv", skip = 11)
dengue_data <- read_csv("raw_data/dengue_data.csv", skip = 11)
gapminder_data <- read.csv("raw_data/gapminder.csv")
## Make tidy. Gapminder doesn't need to be made tidy, just renaming some rows for future use.
flu_tidy <- pivot_longer(data = flu_data, cols = c(2:30), names_to = "Country", values_to = "Value")
flu_tidy <- flu_tidy %>% mutate(Date = str_sub(Date, 1, 4))
flu_tidy <- flu_tidy %>% rename("Year" = "Date")

dengue_tidy <- pivot_longer(data = dengue_data, cols = c(2:11), names_to = "Country", values_to = "Value")
dengue_tidy <- dengue_tidy %>% mutate(Date = str_sub(Date, 1, 4))
dengue_tidy <- dengue_tidy %>% rename("Year" = "Date")

gapminder_data <- as_tibble(gapminder_data)
gapminder_data$year <- as.character(gapminder_data$year)
gapminder_data <- gapminder_data %>% rename("Year" = "year")
gapminder_data <- gapminder_data %>% rename("Country" = "country")
```

An important part of relational data is enabling comparison across different data frame. The data frames are now similar and can be stored into csv and rds files.

```{r export}
# Export data frames as csv and rds
write.csv(flu_tidy, "flu_tidy.csv")
write.csv(dengue_tidy, "dengue_tidy.csv")
write.csv(gapminder_data, "gapminder_tidy.csv")
saveRDS(flu_tidy, "flu_tidy.rds")
saveRDS(dengue_tidy, "dengue_tidy.rds")
saveRDS(gapminder_data, "gapminder_tidy.rds")
```

DBeaver and R were connected (my real password is hidden) and the tables were inserted into the 'workflowsdb' database in DBeaver for inspection.

```{r connecting DBeaver, eval = FALSE}
con <- dbConnect(RPostgres::Postgres(), 
                 dbname = "workflowsdb", 
                 host="localhost", 
                 port="5432", 
                 user="postgres", 
                 password=".") # Hide my real password
```

```{r insert in DBeaver, eval = FALSE}
dbWriteTable(con, "gapminder", gapminder_data)
dbWriteTable(con, "flu", flu_tidy)
dbWriteTable(con, "dengue", dengue_tidy)
```

----

## Inspection {-}

Now that the tables are imported, it is required to check for certain things inside of them. The first thing I checked were the 'NULL' values.

```{sql, connection = con}
SELECT *
  FROM gapminder
WHERE "Year" IS NULL; --make sure no important values are missing
```

```{sql, connection = con}
SELECT *
  FROM gapminder
WHERE "Country" IS NULL;
```

```{sql, connection = con}
SELECT *
  FROM dengue
WHERE "Year" IS NULL;
```

```{sql, connection = con}
SELECT *
  FROM dengue
WHERE "Country" IS NULL;
```

```{sql, connection = con}
SELECT *
  FROM flu
WHERE "Year" IS NULL;
```

```{sql, connection = con}
SELECT *
  FROM flu
WHERE "Country" IS NULL;
```
 
Turns out there were no 'NULL' values in all these tables, which is good. 

Next to checking for 'NULL' values, I also wondered what the total number of cases were and which countries had the highest number of cases. To check this, I used the following scripts.

```{sql, connection = con}
SELECT SUM("Value") --calculate the total number of cases
  FROM dengue;
```

```{sql, connection = con}
SELECT SUM("Value") --calculate the total number of cases
  FROM flu;
```

For a good analysis of the different data they need to be joined together in one table with the important columns. I created a tabel combining all three tables. In my case, I decided to focus on the period between 2002 and 2015. See below for the SQL and R code.

```{r Filter correlating year}
gapminder_filtered <- gapminder_data %>% filter(between(Year, 2002, 2015))
```

```{r Uploading table, eval = FALSE}
dbWriteTable(con, "gapminder_filtered", gapminder_filtered)
```

```{r Combining tables}
flu_dengue_combined <- left_join(flu_tidy, dengue_tidy, by = c("Country", "Year"))
flu_dengue_gapminder_combined <- left_join(flu_dengue_combined, gapminder_filtered, by = c("Country", "Year"))
flu_dengue_gapminder_combined <- flu_dengue_gapminder_combined %>% rename("Flu" = "Value.x")
flu_dengue_gapminder_combined <- flu_dengue_gapminder_combined %>% rename("Dengue" = "Value.y")
```
----

## Visualisations {-}

Now that the dataset had been combined, it is possible to make visualisations of this. In my [self-made Rpackage](https://yourilam.github.io/08_Rpackage.html) I made a function that will help a bit with visualising. With this dataset, the following graphs can be made:

### Life expectancy {-}

```{r visualisation1 life expectancy}
write.csv(flu_dengue_gapminder_combined, "raw_data/package.csv", row.names=FALSE)
life_expectancy_graph("Europe","2002")
```
_Figure 1: Bargraph of the life expectancy for every country in Europe in 2002._

```{r visualisation2 life expectancy}
life_expectancy_graph("Europe","2012")
```
_Figure 2: Bargraph of the life expectancy for every country in Europe in 2015._

###One other visualisation

```{r visualisation3 dengue}
dengue_brazil <- flu_dengue_gapminder_combined %>% filter(Country == "Brazil") %>% select(Dengue, Year)
ggplot(data = dengue_brazil) + geom_col(aes(x = Year, y = Dengue, fill = Year), show.legend = FALSE) + labs(title = "Development of dengue in Brazil each year")
```
_Figure 3: Bargraph of the amount of dengue cases in Brazil every year_
